---
title: "PSTAT 174 Final Project"
author: "Nathan Ho"
date: "2025-03-10"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load in and Explore Data
Original dataset contains Nvidia stock closing prices from 2018 - 2024
New dataset includes Nvidia stock closing prices from January 2022 to September 2024.
```{r}
library(astsa)
library(tseries)
library(timeSeries)
library(forecast)
library(lubridate)
library(dplyr)
library(ggplot2)
library(fGarch)
library(rugarch)
library(pander)

NVDA <- read.csv("~/Downloads/NVIDIA_STOCK.csv") %>%
  rename(Date = Price)

NVDA <- NVDA %>%
  mutate(Date = as.Date(Date, format="%Y-%m-%d")) %>%
  filter(Date >= as.Date("2022-01-03") & Date <= as.Date("2024-09-30")) %>%
  arrange(Date)

ClosingPrice <- NVDA$Close

NVDAts <- ts(as.numeric(ClosingPrice), start = c(2022, 1), frequency = 252)

plot(NVDAts, 
     main="Time Series of NVIDIA (NVDA) Closing Stock Price \n Jan 2022 - Sep 2024", 
     ylab="Closing Price (USD)", 
     xlab="Time")

NVDAts <- as.numeric(NVDAts)
```
Interpretation:
1. Strong Upward Trend:
- Stock price has been increasing over time
- Suggests a long-term growth in Nvidia's stock

2. Exponential Growth Pattern:
- Growth accelerates sharply after mid 2023, suggesting possible non-linear
- Suggests that a log transformation is needed

3. Non-Stationary Data:
- The trend suggests that the data is not stationary, meaning that differencing would be required before fitting an ARIMA model. 
- Exponential increase suggests log transformation is necessary to stabilize variance.

# Check for Stationarity
Use Augmented Dickey-Fuller (ADF) test to test stationarity
```{r}
adf.test(NVDAts)
acf2(NVDAts, 25, main = "ACF/PACF of NVDA Closing Prices")
```
Interpretation:
1. Non-Stationary Data: 
The ADF test p-value = 0.6194 is greater than 0.05, which means that we fail to reject the null hypothesis of stationarity
Meaning: Stock prices exhibit a trend and require differencing before fitting an ARIMA model.
First-order differencing (d=1) is required to remove the trend and make the data stationary.

## Transform
Log Transformation: reduces the heteroskedasticity and helps stabilize variance. Making trends more linear, improving model performance. 
```{r}
NVDALog <- log(NVDAts)
plot(NVDALog, 
     main="Log Time Series of NVIDIA (NVDA) Closing Stock Price", 
     ylab="Closing Price (USD)", 
     xlab="Time", 
     type="l")
```

First order differencing is used to remove trends and make data stationary. NVidia's stock price has a strong long-term trend and differencing is needed to model short term fluctuations effectively.
```{r}
NVDALogDiff = diff(log(NVDAts))
adf.test(NVDALogDiff)
plot(NVDALogDiff, 
     main="First Difference of Log NVIDIA (NVDA) Closing Stock Price", 
     ylab="Closing Price (USD)", 
     xlab="Time", 
     type="l")
acf2(NVDALogDiff, 25, main = "ACF and PACF of First Difference Log Price")
```
p-value = 0.01 reject the null hypothesis (now stationary)
ACF and PACF analysis
Autocorrelation Function (ACF):
- Strong correlation at multiple lags, confirming that there is dependence between stock prices over time
- Gradual decay suggests a moving average (MA) component is necessary
Partial Autocorrelation Function (PACF):
- First lag is significant, indicating the need for autoregressive (AR) terms
Key Insight: Both ACF and PACF indicate the presence of short-term memory effects, meaning SARIMA modeling is appropriate

# Testing different models for ACF/PACFs
```{r}
Sarima1 <- sarima(NVDAts, p=1, d=1, q=1)
summary(arima(NVDAts, order=c(1,1,1)))

Sarima2 <- sarima(NVDAts, p=0, d=1, q=1)
summary(arima(NVDAts, order=c(0,1,1)))

Sarima3 <- sarima(NVDAts, p=1, d=1, q=2)
summary(arima(NVDAts, order=c(1,1,2)))

Sarima4 <- sarima(NVDAts, p=2, d=1, q=1)
summary(arima(NVDAts, order=c(2,1,1)))
```

# Auto arima to confirm (1,1,1)
```{r}
NVDASarima <- arima(NVDAts, order = c(1,1,1))
FcastNVDA <- forecast(NVDASarima, h=252)
plot(FcastNVDA, main="ARIMA(1,1,1) Forecast", 
     ylab="Closing Price (USD)", xlab="Time")
```
Best Model: Arima (1,1,1) with drift

Nvidia's stock price follows a first-order difference ARIMA process with one AR term and one MA term.
Drift term indicates a consistent upward trend in Nvidia stock over time.
Model Residual Diagnostics
- Residuals from ARIMA(1,1,1) with drift:
- Ljung-Box test (p = 0.033) suggests minor residual autocorrelation
- Residuals appear to be randomly distributed.

## Sarima Forecast
```{r}
FcastNVDA <- forecast(NVDASarima, h=252)
plot(FcastNVDA, main="ARIMA(1,1,1) Forecast", 
     ylab="Closing Price (USD)", xlab="Time")
```
252 day (1-year) forecast using SARIMA (1,1,1) with drift.
Forecasted values show continued growth.
Confidence Intervals (80%, 95%) widen over time, meaning uncertainty increases as we project further.
Nvidia's stock is projected to continue its strong growth trend.
Potential volatility is reflected in the widening confidence bands.

## Garch
```{r}
NVDAarma <- arima(NVDALogDiff, order = c(1,1,1))
residuals <- residuals(NVDAarma)
squaredresiduals = residuals^2
acf2(squaredresiduals, 25, main = "ACF/PACF of Squared Residuals")
```

```{r}
library(fGarch)

NVDAGarch1 <- garchFit(~ arma(1,1) + garch(1,1), data=NVDALogDiff, cond.dist="std")

NVDAGarch2 <- garchFit(~ arma(1,1) + garch(1,0), data=NVDALogDiff, cond.dist="std")

NVDAGarch3 <- garchFit(~ arma(0,0) + garch(1,0), data=NVDALogDiff, cond.dist="std")

summary(NVDAGarch1)
summary(NVDAGarch2)
summary(NVDAGarch3)
```

```{r}
par(mfrow=c(1,2))
plot(NVDAGarch1, which = c(1,3,9,10))
plot(NVDAGarch1, which =13)
```

```{r}
ForecastGARCH <- predict(NVDAGarch1, n.ahead = 252, plot=TRUE)
summary(ForecastGARCH)
```

